\pkg{partykit}: \\ A Toolbox for Recursive Partytioning

### 2009-07-02 pkg/inst/slides/useR-2009_slides.Rnw
Z comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@137 255f4249-2263-4417-b1ea-8498779e27a7

\author{Torsten Hothorn \\ Ludwig-Maximilians-Universit\"at M\"unchen \vspace*{1cm} \\ 
        Achim Zeileis \\ Wirtschaftsuniversit\"at Wien}
\fancyfoot[L]{\tiny 2009-07-09} %%\copyright{} 2007 Friedrich Leisch
\fancyfoot[R]{\arabic{page}}

@@

The Machine Learning task view lists the following tree-related packages
\begin{compactitem}
\item \pkg{rpart} (CART)
\item \pkg{tree} (CART)
\item \pkg{mvpart} (multivariate CART)
\item \pkg{knnTree} (nearest neighbor based trees)
\item \pkg{RWeka} (J48, M5$^\prime$, LMT)
\item \pkg{LogicReg} (Logic Regression)
\item \pkg{BayesTree} (with Bayesian flavor)
\item \pkg{TWIX} (with extra splits)
\item \pkg{party} (conditional inference trees, model-based trees)
\item ...
\end{compactitem}

@@

Moreover, important methods in this field, mostly from the statistics literature,
are missing from the above list, for example
\begin{compactitem}
\item CHAID
\item FACT
\item QUEST
\item GUIDE
\item LOTUS
\item C4.5
\item ...
\end{compactitem}

@@

... have a common infrastructure for
\begin{itemize}
\item representing fitted trees,
\item printing trees,
\item plotting trees, and
\item computing predictions from trees?
\end{itemize}

@@

offers a unified representation of trees with methods for
\begin{itemize}
\item printing,
\item plotting,
\item computing predictions and
\item coercing from/to different formats.
\end{itemize}

@@

<<ex-rpart, echo = TRUE>>=
data("GlaucomaM", package = "ipred")
g_rpart <- rpart(Class ~ ., data = GlaucomaM)
print(g_rpart)
@

@@

{\small
<<ex-rpart-partykit, echo = TRUE>>=
g_party <- as.party(g_rpart)
print(g_party, header = FALSE)
@
}

@@

The resulting package \pkg{CHAID} is available from R-forge and
consists of only 362 lines of R code and was validated against SPSS 15.

@@

\small{
<<ex-CHAID, echo = TRUE>>=
data("BreastCancer", package = "mlbench")
b_chaid <- chaid(Class ~ Cl.thickness + Cell.size + Cell.shape + 
                  Marg.adhesion + Epith.c.size + Bare.nuclei + 
                  Bl.cromatin + Normal.nucleoli + Mitoses, 
                  data = BreastCancer)
@
}

@@

Benchmark experiment CHAID vs. CTREE\\
(Bachelor thesis by Stefanie Thiemichen)
\begin{center}
\includegraphics{ratiobox}
\end{center}

@@

<<ex-pmml-predict, echo = TRUE>>=
nd <- data.frame(nutzung = gl(3, 10), beruf = sample(gl(3, 10)))
predict(pm, newdata = nd)
@

@@

Once a stable version of \pkg{partykit} is available from CRAN, we hope that
useRs will start
\begin{itemize}
\item adding new tree algorithms,
\item adding coercing methods from/to other packages,
\item adding new plot functionality, and generally
\item will use our kit to party harder.
\end{itemize}

>>>>

\author{Torsten Hothorn \\ Ludwig-Maximilians-Universit\"at M\"unchen \vspace*{1cm} \\ 
        Achim Zeileis \\ WU Wirtschaftsuniversit\"at Wien}
\fancyfoot[L]{\tiny 2009-07-09} %%\copyright{} 2007 Friedrich Leisch
\fancyfoot[R]{\arabic{page}}

@@

The Machine Learning task view lists the following tree-related packages
\begin{compactitem}
\item \pkg{rpart} (CART)
\item \pkg{tree} (CART)
\item \pkg{mvpart} (multivariate CART)
\item \pkg{knnTree} (nearest-neighbor-based trees)
\item \pkg{RWeka} (J4.8, M5$^\prime$, LMT)
\item \pkg{LogicReg} (Logic Regression)
\item \pkg{BayesTree} (with Bayesian flavor)
\item \pkg{TWIX} (with extra splits)
\item \pkg{party} (conditional inference trees, model-based trees)
\item ...
\end{compactitem}

@@

Moreover, important methods in this field, mostly from the statistics literature,
are missing from the above list, for example
\begin{compactitem}
\item CHAID
\item FACT
\item QUEST
\item GUIDE
\item LOTUS
\item CRUISE
\item ...
\end{compactitem}

@@

{\color{white} hansi \\}
... have a common infrastructure for
\begin{itemize}
\item representing fitted trees,
\item printing trees,
\item plotting trees, and
\item computing predictions from trees?
\end{itemize}

@@

{\color{white} hansi \\}
offers a unified approach to 
\begin{itemize}
\item representing fitted trees,
\item printing trees,
\item plotting trees, and
\item computing predictions from trees.
\end{itemize}

@@

<<ex-rpart, echo = TRUE>>=
data("GlaucomaM", package = "ipred")
library("rpart")
g_rpart <- rpart(Class ~ ., data = GlaucomaM)
print(g_rpart)
@

@@

{\small
<<ex-rpart-partykit, echo = TRUE>>=
library("partykit")
g_party <- as.party(g_rpart)
print(g_party, header = FALSE)
@
}

@@

The resulting package \pkg{CHAID} is available from R-Forge and
consists of only 362 lines of R code and was validated against SPSS 15.

@@

\small{
<<ex-CHAID, echo = TRUE>>=
data("BreastCancer", package = "mlbench")
library("CHAID")
b_chaid <- chaid(Class ~ Cl.thickness + Cell.size + Cell.shape + 
                  Marg.adhesion + Epith.c.size + Bare.nuclei + 
                  Bl.cromatin + Normal.nucleoli + Mitoses, 
                  data = BreastCancer)
@
}

@@

Benchmark experiment misclassification error CHAID vs.~CTREE\\
(Bachelor thesis by Stefanie Thiemichen)
\begin{center}
\includegraphics{ratiobox}
\end{center}

@@

<<ex-pmml-predict, echo = TRUE>>=
nd <- data.frame(X1 = gl(3, 10), X2 = sample(gl(3, 10)))
predict(pm, newdata = nd)
@

@@

Once a stable version of \pkg{partykit} is available from CRAN, we hope that
useRs will start
\begin{itemize}
\item adding new tree algorithms,
\item adding new ensemble methods,
\item adding coercing methods from/to other packages,
\item adding new plot functionality, and generally
\item will use our kit to party harder.
\end{itemize}
